---
title: "Plots for Global Cost of Cash Benchmarking Deck"
author: "Ben Mazzotta"
date: "Friday, March 27, 2015"
output: html_document
---

This document creates plots from the Global Cost of Cash Benchmarking Study. The plots are written for HD dimensions:

* 1280px wide
* 720px tall
* 18pt font


```{r prelims}
require(ggplot2); require(data.table)
setwd("../data")
load("ccglobal.plot.Rda")


attr(costofcash$region, "levels")
setattr(costofcash$region, "levels", c("Aggregates","Asia Pac","Eur Asia", "Latin Am Car","Mideast NAfr", "North Amer","South Asia","SS Africa"))
costofcash[,income:=ordered(income, c("High income: OECD", "High income: nonOECD", "Upper middle income", "Lower middle income","Low income","Aggregates","Not classified"))]

```

## Warm-ups

These charts validate the graphic design in `markdown` and `ggplot2`. Once we like the design, it should be plug and play.

```{r plotstart, echo=FALSE, fig.width=9, fig.height=5, dpi=144, fig.path="../images/"}
# boxplot(pop_MM~region, data=costofcash, horizontal=TRUE, las=1, log="x")
qplot(region, pop_MM, data=costofcash, geom="boxplot", log="y")+coord_flip()

plot1 <- ggplot(costofcash[!is.na(region),], aes(region, log(pop_MM)))
plot1 + geom_boxplot(aes(fill=income))+coord_flip()+scale_fill_brewer()

plot1+geom_jitter(aes(color=income))+scale_color_brewer(palette="Set1")+coord_flip()
```


## ATM Costs


Charts planned: 

* Regional boxplots for national costs
* Regional boxplots for unit costs
* Scatterplots for national vs. unit costs, by region

```{r  plotATM, echo=FALSE, fig.width=9, fig.height=5, dpi=144, fig.path="../images/"}


plot1 <- ggplot(costofcash[!is.na(region)], aes(region, log(ATMUnitCost)))
plot2 <- ggplot(costofcash[!is.na(region)], aes(region, log(ATMNatlCost)))
plot3 <- ggplot(costofcash, aes(log(ATMNatlCost), log(ATMUnitCost)))

plot1 + geom_boxplot(aes(fill=region))+coord_flip()+scale_fill_brewer(palette="Blues") + 
  labs(title="Unit Costs of ATM Maintenance \n Regional Comparison", y="ATM Unit Cost (log scale, USD)", x=NULL)
plot2 + geom_boxplot(aes(fill=region))+coord_flip()+scale_fill_brewer(palette="Greens")+ 
  labs(title="National Fleet Costs of ATM Maintenance \n Regional Comparison", y="ATM Fleet Cost (log scale, USD)", x=NULL)
plot3 + geom_point(aes(color=region)) + scale_colour_brewer(palette="Set1")+ 
  labs(title="ATM Fleet Costs vs. Unit Costs \n Global Scatterplot by Region", x="ATM Fleet Cost (log scale, USD)", y="ATM Unit Cost (log scale, USD)")


##    Drill down by income
plot1 + geom_boxplot(aes(fill=income))+coord_flip()+scale_fill_brewer(palette="Blues")+ 
  labs(title="Unit Costs of ATM Maintenance \n Regional Comparison, Drill Down by Income", y="ATM Unit Cost (log scale, USD)", x=NULL)
plot2 + geom_boxplot(aes(fill=income))+coord_flip()+scale_fill_brewer(palette="Greens")+ 
  labs(title="National Fleet Costs of ATM Maintenance \n Regional Comparison, Drill Down by Income", y="ATM Fleet Cost (log scale, USD)", x=NULL)
plot3 + geom_point(aes(colour=income)) + scale_colour_brewer(palette="Set1")+ 
  labs(title="ATM Fleet Costs vs. Unit Costs \n Global Scatterplot by Income", x="ATM Fleet Cost (log scale, USD)", y="ATM Unit Cost (log scale, USD)")


qplot(log(ATMNatlCost), log(ATMUnitCost), data=costofcash[region=="SS Africa"], geom="text", label=iso2c, color=income)+scale_color_brewer(palette="Dark2") + 
  labs(title="National ATM Fleet Costs \n Regional Focus: Africa", x="National Cost of ATMs (log scale, USD)", y="Unit Cost of ATMs (log scale, USD)")
qplot(log(ATMNatlCost), log(ATMUnitCost), data=costofcash[region=="Eur Asia"], geom="text", label=iso2c, color=income)+scale_color_brewer(palette="Dark2")+ 
  labs(title="National ATM Fleet Costs \n Regional Focus: Europe and Central Asia", x="National Cost of ATMs (log scale, USD)", y="Unit Cost of ATMs (log scale, USD)")
qplot(log(ATMNatlCost), log(ATMUnitCost), data=costofcash[region=="South Asia"], geom="text", label=iso2c, color=income)+scale_color_brewer(palette="Dark2")+ 
  labs(title="National ATM Fleet Costs \n Regional Focus: South Asia", x="National Cost of ATMs (log scale, USD)", y="Unit Cost of ATMs (log scale, USD)")
# qplot(log(ATMNatlCost), log(ATMUnitCost), data=costofcash[region=="North Amer"], geom="text", label=iso2c, color=income)+scale_color_brewer(palette="Dark2")+ 
#   labs(title="National ATM Fleet Costs \n Regional Focus: North America", x="National Cost of ATMs (log scale, USD)", y="Unit Cost of ATMs (log scale, USD)")
qplot(log(ATMNatlCost), log(ATMUnitCost), data=costofcash[region=="Latin Am Car"], geom="text", label=iso2c, color=income)+scale_color_brewer(palette="Dark2")+ 
  labs(title="National ATM Fleet Costs \n Regional Focus: Latin America and Caribbean", x="National Cost of ATMs (log scale, USD)", y="Unit Cost of ATMs (log scale, USD)")
qplot(log(ATMNatlCost), log(ATMUnitCost), data=costofcash[region=="Asia Pac"], geom="text", label=iso2c, color=income)+scale_color_brewer(palette="Dark2")+ 
  labs(title="National ATM Fleet Costs \n Regional Focus: Asia Pacific", x="National Cost of ATMs (log scale, USD)", y="Unit Cost of ATMs (log scale, USD)")
qplot(log(ATMNatlCost), log(ATMUnitCost), data=costofcash[region=="Mideast NAfr"], geom="text", label=iso2c, color=income)+scale_color_brewer(palette="Dark2")+ 
  labs(title="National ATM Fleet Costs \n Regional Focus: Mideast and North Africa", x="National Cost of ATMs (log scale, USD)", y="Unit Cost of ATMs (log scale, USD)")


```

## Cash in transit

Planned charts

* World map
* Regional rankings
* Scatterplots of ATM vs CIT
* Scatterplots of CIT vs Finclusion


```{r plotcashintransit, fig.width=9, fig.height=5, dpi=144, fig.path="../images/"}
plot1 <- ggplot(costofcash[!is.na(region)], aes(region, log(CITest)))
plot2 <- ggplot(costofcash[!is.na(income)], aes(income, log(CITest)))
plot3 <- ggplot(costofcash, aes(log(ATMNatlCost), log(CITest)))


## Base plots
plot1 + geom_boxplot() + coord_flip()
plot2 + geom_boxplot() + coord_flip() 
plot3 + geom_point() 

## Drill down 
plot1 + geom_boxplot(aes(fill=income)) + coord_flip() + scale_fill_brewer(palette="Set1") 
plot2 + geom_boxplot(aes(fill=region)) + coord_flip() + scale_fill_brewer(palette="Dark2") 
plot3 + geom_point(aes(color=income))  + scale_color_brewer(palette="Set1") 
plot3 + geom_point(aes(color=region)) + scale_color_brewer(palette="Dark2") 


##    Regional Scatterplots
qplot(ATMNatlCost, CITest, data=costofcash[region=="SS Africa"], geom="text", color=income, label=iso2c, log="xy") + scale_color_brewer(palette="Dark2") + 
  labs(title="CIT Costs in Africa (Modeled) \n Regional Focus and Comparison to ATM Fleet Cost", x="ATM Fleet Costs (log scale USD)", y="CIT Modeled Fleet Costs (log scale USD)")

qplot(ATMNatlCost, CITest, data=costofcash[region=="South Asia"], geom="text", color=income, label=iso2c, log="xy") + scale_color_brewer(palette="Dark2") + 
  labs(title="CIT Costs in South Asia (Modeled) \n Regional Focus and Comparison to ATM Fleet Cost", x="ATM Fleet Costs (log scale USD)", y="CIT Modeled Fleet Costs (log scale USD)")

qplot(ATMNatlCost, CITest, data=costofcash[region=="Asia Pac"], geom="text", color=income, label=iso2c, log="xy") + scale_color_brewer(palette="Dark2") + 
  labs(title="CIT Costs in Asia Pacific Region (Modeled) \n Regional Focus and Comparison to ATM Fleet Cost", x="ATM Fleet Costs (log scale USD)", y="CIT Modeled Fleet Costs (log scale USD)")

qplot(ATMNatlCost, CITest, data=costofcash[region=="Mideast NAfr"], geom="text", color=income, label=iso2c, log="xy") + scale_color_brewer(palette="Dark2") + 
  labs(title="CIT Costs in Mideast and North Africa (Modeled) \n Regional Focus and Comparison to ATM Fleet Cost", x="ATM Fleet Costs (log scale USD)", y="CIT Modeled Fleet Costs (log scale USD)")

qplot(ATMNatlCost, CITest, data=costofcash[region=="Latin Am Car"], geom="text", color=income, label=iso2c, log="xy") + scale_color_brewer(palette="Dark2") + 
  labs(title="CIT Costs in Latin America and Caribbean (Modeled) \n Regional Focus and Comparison to ATM Fleet Cost", x="ATM Fleet Costs (log scale USD)", y="CIT Modeled Fleet Costs (log scale USD)")


## Income Group Focus
qplot(ATMNatlCost, CITest, data=costofcash[income=="High income: nonOECD"], geom="text", color=income, label=iso2c, log="xy") + scale_color_brewer(palette="Dark2") + 
  labs(title="CIT Costs in Africa (Modeled) \n Regional Focus and Comparison to ATM Fleet Cost", x="ATM Fleet Costs (log scale USD)", y="CIT Modeled Fleet Costs (log scale USD)")


## Regional focus
plot1 + geom_boxplot() + coord_flip() + facet_wrap(~income)
plot2 + geom_boxplot() + coord_flip() + facet_wrap(~region)
plot3 + geom_point(aes(color=income))  + scale_color_brewer(palette="Set1") + facet_wrap(~region)
plot3 + geom_point(aes(color=region))  + scale_color_brewer(palette="Dark2") + facet_wrap(~income)

```



## Consumer costs

Planned charts

* World map
* regional rankings
* Scatterplot of time vs. fees
  * Regional focus 
  * Income bracket focus
* Scatterplot of time vs financial inclusion


```{r plothouseholdcosts, fig.width=9, fig.height=5, dpi=144, fig.path="../images/"}
## Hello
plot1 <- ggplot(costofcash, aes(log10(cashfees), log10(transcost)))
plot2 <- ggplot(costofcash, aes(log10(cashfees), log10(timespent)))
plot3 <- ggplot(costofcash, aes(log10(transcost), log10(timespent)))
plot4 <- ggplot(costofcash, aes(log10(transcost/pop_MM), log10(timespent/pop_MM)))


##    Facets by region
plot1 + geom_point() + facet_wrap(~region) + 
  labs(title="Cost of Cash: Fees vs. Transit", x="Cash Fees (log scale, USD MM)", y="Transit Cost (log scale, USD MM)")
plot2 + geom_point() + facet_wrap(~region)+ 
  labs(title="Cost of Cash: Fees vs. Time Spent", x="Cash Fees (log scale, USD MM)", y="Time spent (log scale, minutes MM)")
plot3 + geom_point() + facet_wrap(~region)+ 
  labs(title="Cost of Cash: Transit vs. Time Spent", x="Cash Fees (log scale, USD MM)", y="Time spent (log scale, minutes MM)")
plot4 + geom_point() + facet_wrap(~region)+ 
  labs(title="Cost of Cash: Per capita transit vs. Per capita time spent", x="Transit cost per capita (log scale, USD MM)", y="Time spent per capita (log scale, minutes)")


plot4 + geom_point(aes(color=region)) + scale_color_brewer(palette="Set1") + 
  labs(title="Cost of Cash: Per capita transit vs. Per capita time spent", x="Transit cost per capita (log scale, USD MM)", y="Time spent per capita (log scale, minutes)")


## Color scatterplots by income

plot1 + geom_point(aes(color=income)) + scale_color_brewer(palette="RdBu") + 
  labs(title="Cost of Cash: Fees vs. Transit", x="Cash Fees (log scale, USD MM)", y="Transit Cost (log scale, USD MM)")
plot2 + geom_point(aes(color=income)) + scale_color_brewer(palette="RdBu") + 
  labs(title="Cost of Cash: Fees vs. Time Spent", x="Cash Fees (log scale, USD MM)", y="Time spent (log scale, minutes MM)")
plot3 + geom_point(aes(color=income)) + scale_color_brewer(palette="RdBu") + 
  labs(title="Cost of Cash: Transit vs. Time Spent", x="Cash Fees (log scale, USD MM)", y="Time spent (log scale, minutes MM)")

plot4 + geom_point(aes(color=income)) + scale_color_brewer(palette="RdBu") + 
  labs(title="Cost of Cash: Per capita transit vs. Per capita time spent", x="Transit cost per capita (log scale, USD MM)", y="Time spent per capita (log scale, minutes)")


##    Focus by region
qplot(log10(transcost), log10(cashfees), data=costofcash[region=="SS Africa"], geom="text", label=iso2c, color=income) + scale_color_brewer(palette="Set1")+
  labs(title="Household Cash Costs: Cash Fees and Transit Costs \n Regional Focus: Sub-Saharan Africa", x="Transit Cost (log scale, USD MM)", y="Cash Fees (log scale, USD MM)")

qplot(log10(transcost), log10(cashfees), data=costofcash[region=="Asia Pac"], geom="text", label=iso2c, color=income) + scale_color_brewer(palette="Set1")+
  labs(title="Household Cash Costs: Cash Fees and Transit Costs \n Regional Focus: East Asia Pacific", x="Transit Cost (log scale, USD MM)", y="Cash Fees (log scale, USD MM)")

qplot(log10(transcost), log10(cashfees), data=costofcash[region=="Eur Asia"], geom="text", label=iso2c, color=income) + scale_color_brewer(palette="Set1")+
  labs(title="Household Cash Costs: Cash Fees and Transit Costs \n Regional Focus: Europe and Central Asia", x="Transit Cost (log scale, USD MM)", y="Cash Fees (log scale, USD MM)")

qplot(log10(transcost), log10(cashfees), data=costofcash[region=="Latin Am Car"], geom="text", label=iso2c, color=income) + scale_color_brewer(palette="Set1")+
  labs(title="Household Cash Costs: Cash Fees and Transit Costs \n Regional Focus: Latin America and Caribbean", x="Transit Cost (log scale, USD MM)", y="Cash Fees (log scale, USD MM)")

qplot(log10(transcost), log10(cashfees), data=costofcash[region=="Mideast NAfr"], geom="text", label=iso2c, color=income) + scale_color_brewer(palette="Set1")+
  labs(title="Household Cash Costs: Cash Fees and Transit Costs \n Regional Focus: Mideast and North Africa", x="Transit Cost (log scale, USD MM)", y="Cash Fees (log scale, USD MM)")

qplot(log10(transcost), log10(cashfees), data=costofcash[region=="South Asia"], geom="text", label=iso2c, color=income) + scale_color_brewer(palette="Set1")+
  labs(title="Household Cash Costs: Cash Fees and Transit Costs \n Regional Focus: South Asia", x="Transit Cost (log scale, USD MM)", y="Cash Fees (log scale, USD MM)")


```


## Currency issuance

Planned charts

* Cost of new issuance
  * per capita costs vs. per capita GDP
  * per capita costs vs. financial inclusion
* Value of seigniorage
  * per capita costs vs. per capita GDP
  * per capita costs vs. financial inclusion
* Scatterplots of seigniorage vs. issuance
  * national costs
  * per capita costs
  * Regional focus
  * income group focus
  
```{r plotcurrencyissuance, fig.width=9, fig.height=5, dpi=144, fig.path="../images/"}
## Value of seigniorage

plot1 <- ggplot(costofcash, aes(log10(gdp*10^-6), log10(seigniorage)))
plot2 <- ggplot(costofcash[!is.na(region),], aes(region, log10(seigniorage)))
plot3 <- ggplot(costofcash[!is.na(income),], aes(income, log10(seigniorage)))

plot1 + geom_text(aes(color=region, label=iso2c)) + scale_color_brewer(palette="Dark2")
plot1 + geom_point(aes(color=region)) + scale_color_brewer(palette="Dark2")
plot2 + geom_boxplot(aes(color=region)) + scale_color_brewer(palette="Dark2") + coord_flip()
plot3 + geom_boxplot(aes(color=income)) + scale_color_brewer(palette="Dark2") + coord_flip()

plot11 <- ggplot(costofcash[region=="SS Africa" & !is.na(seigniorage),], aes(log10(gdp*10^-6), log10(seigniorage), label=iso2c, color=income))

plot12 <- ggplot(costofcash[region=="South Asia" & !is.na(seigniorage),], aes(log10(gdp*10^-6), log10(seigniorage), label=iso2c, color=income))

plot13 <- ggplot(costofcash[region=="Mideast NAfr" & !is.na(seigniorage),], aes(log10(gdp*10^-6), log10(seigniorage), label=iso2c, color=income))

plot14 <- ggplot(costofcash[region=="Latin Am Car" & !is.na(seigniorage),], aes(log10(gdp*10^-6), log10(seigniorage), label=iso2c, color=income))

plot15 <- ggplot(costofcash[region=="Eur Asia" & !is.na(seigniorage),], aes(log10(gdp*10^-6), log10(seigniorage), label=iso2c, color=income))

plot16 <- ggplot(costofcash[region=="Asia Pac" & !is.na(seigniorage),], aes(log10(gdp*10^-6), log10(seigniorage), label=iso2c, color=income))



plot11 + geom_text() + 
  scale_color_brewer(palette="Set1") + labs(title="Seigniorage Income Vs. GDP \n Regional Focus: Africa", x="GDP at PPP (log scale, USD MM)", y="Seigniorage (log scale, USD MM)")
plot12 + geom_text() + 
  scale_color_brewer(palette="Set1") + labs(title="Seigniorage Income Vs. GDP \n Regional Focus: South Asia", x="GDP at PPP (log scale, USD MM)", y="Seigniorage (log scale, USD MM)")
plot13 + geom_text() + 
  scale_color_brewer(palette="Set1") + labs(title="Seigniorage Income Vs. GDP \n Regional Focus: Mideast North Africa", x="GDP at PPP (log scale, USD MM)", y="Seigniorage (log scale, USD MM)")
plot14 + geom_text() + 
  scale_color_brewer(palette="Set1") + labs(title="Seigniorage Income Vs. GDP \n Regional Focus: Latin America Caribbean", x="GDP at PPP (log scale, USD MM)", y="Seigniorage (log scale, USD MM)")
plot15 + geom_text() + 
  scale_color_brewer(palette="Set1") + labs(title="Seigniorage Income Vs. GDP \n Regional Focus: Europe and Central Asia", x="GDP at PPP (log scale, USD MM)", y="Seigniorage (log scale, USD MM)")
plot16 + geom_text() + 
  scale_color_brewer(palette="Set1") + labs(title="Seigniorage Income Vs. GDP \n Regional Focus: East Asia Pacific", x="GDP at PPP (log scale, USD MM)", y="Seigniorage (log scale, USD MM)")

##  Cost of Currency Issuance

plot1 <- ggplot(costofcash, aes(log10(gdp*10^-6), log10(CurrIssue))) 
plot2 <- ggplot(costofcash[!is.na(region),], aes(region, log10(CurrIssue)))
plot3 <- ggplot(costofcash[!is.na(income),], aes(income, log10(CurrIssue)))


plot1 + geom_text(aes(color=region, label=iso2c)) + scale_color_brewer(palette="Dark2") + labs(title="Currency Issuance", x="GDP (log scale, USD MM)", y="Currency Issuance Costs (log scale, USD MM)")

plot1 + geom_point(aes(color=region)) + scale_color_brewer(palette="Dark2")+ labs(title="Currency Issuance", x="GDP (log scale, USD MM)", y="Currency Issuance Costs (log scale, USD MM)")

plot2 + geom_boxplot(aes(color=region)) + scale_color_brewer(palette="Dark2") + coord_flip()+ labs(title="Curreny Issuance \n Boxplots by Region", x=NULL, y="Currency Issuance Costs (log scale, USD MM)")

plot3 + geom_boxplot(aes(color=income)) + scale_color_brewer(palette="Dark2") + coord_flip()+ labs(title="Currency Issuance \n Boxplots by Income Group", x=NULL, y="Currency Issuance Costs (log scale, USD MM)")
 

## Currency issuance vs. seigniorage revenue

plot1 <- ggplot(costofcash, aes(log10(CurrIssue), log10(seigniorage), color=income, label=iso2c)) + labs(title="Currency Issuance vs. Seigniorage Revenue", x="Currency Issuance (log scale, USD MM)", y="Seigniorage Revenue (log scale, USD MM)")

plot11 <- ggplot(costofcash[region=="SS Africa"], aes(log10(CurrIssue), log10(seigniorage), color=income, label=iso2c)) + labs(title="Currency Issuance vs. Seigniorage Revenue \n Regional Focus: Africa", x="Currency Issuance (log scale, USD MM)", y="Seigniorage Revenue (log scale, USD MM)")

plot12 <- ggplot(costofcash[region=="South Asia"], aes(log10(CurrIssue), log10(seigniorage), color=income, label=iso2c)) + labs(title="Currency Issuance vs. Seigniorage Revenue \n Regional Focus: South Asia", x="Currency Issuance (log scale, USD MM)", y="Seigniorage Revenue (log scale, USD MM)")

plot13 <- ggplot(costofcash[region=="Mideast NAfr"], aes(log10(CurrIssue), log10(seigniorage), color=income, label=iso2c)) + labs(title="Currency Issuance vs. Seigniorage Revenue \n Regional Focus: Mideast and North Africa", x="Currency Issuance (log scale, USD MM)", y="Seigniorage Revenue (log scale, USD MM)")

plot14 <- ggplot(costofcash[region=="Asia Pac"], aes(log10(CurrIssue), log10(seigniorage), color=income, label=iso2c)) + labs(title="Currency Issuance vs. Seigniorage Revenue \n Regional Focus: East Asia Pacific", x="Currency Issuance (log scale, USD MM)", y="Seigniorage Revenue (log scale, USD MM)")

plot15 <- ggplot(costofcash[region=="Latin Am Car"], aes(log10(CurrIssue), log10(seigniorage), color=income, label=iso2c)) + labs(title="Currency Issuance vs. Seigniorage Revenue \n Regional Focus: Latin America Caribbean", x="Currency Issuance (log scale, USD MM)", y="Seigniorage Revenue (log scale, USD MM)")


plot16 <- ggplot(costofcash[region=="Eur Asia"], aes(log10(CurrIssue), log10(seigniorage), color=income, label=iso2c)) + labs(title="Currency Issuance vs. Seigniorage Revenue \n Regional Focus: Europe Central Asia", x="Currency Issuance (log scale, USD MM)", y="Seigniorage Revenue (log scale, USD MM)")




```

  
## Tax gap

Planned charts

* Value of tax gap
  * Regional comparison
  * Income group comparison
  * Scatterplot vs. per capita GDP
* Scatterplots of seigniorage vs. tax gap
  * national costs
  * regional focus
  * income group focus
* Scatterplots of currency issuance vs. tax gap
  * national costs
  * regional focus
  * income group focus

  
```{r plottaxgap, fig.width=9, fig.height=5, dpi=144, fig.path="../images/"}
## Hello

plot1 <- ggplot(costofcash, aes(seigniorage/gdp*10^8, taxgap, color=income, label=iso2c)) 
plot2 <- ggplot(costofcash[region=="SS Africa",], aes(region, taxgap, color=income))
plot3 <- ggplot(costofcash[!is.na(income),], aes(income, taxgap))
plot4 <- ggplot(costofcash, aes(GDPcap, taxgap, color=income, label=iso2c))


plot11 <- ggplot(costofcash[region=="SS Africa",], aes(seigniorage/gdp*10*8, taxgap, color=income, label=iso2c))
plot12 <- ggplot(costofcash[region=="South Asia",], aes(seigniorage/gdp*10*8, taxgap, color=income, label=iso2c))
plot13 <- ggplot(costofcash[region=="Mideast NAfr",], aes(seigniorage/gdp*10*8, taxgap, color=income, label=iso2c))
plot14 <- ggplot(costofcash[region=="Latin Am Car",], aes(seigniorage/gdp*10*8, taxgap, color=income, label=iso2c))
plot15 <- ggplot(costofcash[region=="Eur Asia",], aes(seigniorage/gdp*10*8, taxgap, color=income, label=iso2c))
plot16 <- ggplot(costofcash[region=="Asia Pac",], aes(seigniorage/gdp*10*8, taxgap, color=income, label=iso2c))

plot1 + geom_point() + scale_color_brewer(palette="RdBu") + labs(title="Tax Gap vs. Seigniorage Revenue", x="Seigniorage (% of GDP)", y="Tax Gap (% of GDP)")

plot1 + geom_text() + scale_color_brewer(palette="RdBu") + labs(title="Tax Gap vs. Seigniorage Revenue", x="Seigniorage (% of GDP)", y="Tax Gap (% of GDP)")


plot11 + geom_text() + scale_color_brewer(palette="RdBu") + labs(title="Tax Gap \n Regional Focus: Africa", x="Seigniorage (% of GDP)", y="Tax Gap (% of GDP)")

plot12 + geom_text() + scale_color_brewer(palette="RdBu") + labs(title="Tax Gap \n Regional Focus: South Asia", x="Seigniorage (% of GDP)", y="Tax Gap (% of GDP)")

plot13 + geom_text() + scale_color_brewer(palette="RdBu")  + labs(title="Tax Gap \n Regional Focus: Mideast North Africa", x="Seigniorage (% of GDP)", y="Tax Gap (% of GDP)")

plot14 + geom_text() + scale_color_brewer(palette="RdBu")  + labs(title="Tax Gap \n Regional Focus: Latin America Caribbean", x="Seigniorage (% of GDP)", y="Tax Gap (% of GDP)")

plot15 + geom_text() + scale_color_brewer(palette="RdBu")  + labs(title="Tax Gap \n Regional Focus: Europe Central Asia", x="Seigniorage (% of GDP)", y="Tax Gap (% of GDP)")

plot16 + geom_text() + scale_color_brewer(palette="RdBu")  + labs(title="Tax Gap \n Regional Focus: East Asia Pacific", x="Seigniorage (% of GDP)", y="Tax Gap (% of GDP)")

```

## Cash hoards

Data is not yet ready.