---
title: "Cash fees multiple imputation 2"
author: "Ben Mazzotta"
date: "Thursday, October 09, 2014"
output: html_document
---

Second try to mark down the cash fees data.

## Load the data

```{r prelims, warning=FALSE, echo=FALSE,results='markup', message=FALSE}
require(ggplot2) ; require(Amelia) ; require(data.table) ; require(Zelig)
setwd("../data")
load("working.Rdata")
# names(node3)
# summary(node3)
# tables()
setnames(node3, "acc_active", "finclactive")
setnames(node3, "rec_paymt", "finclpay")
setnames(node3, "rec_wage", "finclwage")
setnames(node3, "rec_remit", "finclremit")
setnames(node3, "remit_MM", "remitValue")
setnames(node3, "remit_TRX", "remitTrx")

cat("Ready for imputation.")
```

## Multiple imputation

The first step of the imputations is to conduct multiple imputation using as many covariates as possible.

* GDPcap
* acc_active
* rec_payment
* rec_wage
* remit_TRX
* cashprice
* country


```{r impute}
imp01 <- amelia(x=node3, cs="country", id=c("year","iso3c","income","region","lending", "wb3c","payXrem","iso2c"), log=c("gdpcap","pop","remitValue","remitTrx","gdpreal","gdp"), logst=c("finclactive","finclwage","fincpay","finclremit"), empri=1)
summary(imp01)
names(imp01)
```


## Fit

Step #2 is to fit the linear model. 

```{r fit_zelig}

mod01 <- zelig(formula="cashprice~gdpcap+finclactive+remitTrx", model="ls", data = imp01)
mod02 <- zelig(formula="cashprice~gdpcap+finclactive+remitTrx", model="ls", data = node3)
summary(mod01) ;summary(mod02)
names(mod01)

```

## Predict

Step #3 is to predict from the linear model. 

```{r sim_zelig, fig.height=4.5, fig.width=8}

##     setx()  command tells Zelig that the input values are the imputed datasets.
# str(as.matrix(node3[,.SD,.SDcols=c("gdpcap","finclactive","remitTrx")]))
pred01x <- setx(mod01, fn=NULL)

##    sim() command produces estimates from the five estimates of the linear model.
pred01 <- sim(mod01, x=pred01x)

##    Regrettably they are stored separately.
summary(pred01)
names(pred01[[1]]$result)

##    But you can pull them out one at a time.
price01<- pred01[[1]]$result$fitted.values
summary(price01)
pricehat01 <- imp01$imputations$imp1$cashprice
summary(pricehat01)
# priceplot01 <- data.frame(price01, pricehat01)
# qplot(price01, pricehat01, data=priceplot01, geom="point")


price02 <- pred01[[2]]$result$fitted.values
summary(price02)
price03 <- pred01[[3]]$result$fitted.values
summary(price03)
price04 <- pred01[[4]]$result$fitted.values
summary(price04)
price05 <- pred01[[5]]$result$fitted.values
summary(price05)

##      And append them into a data frame, calculate the mean.
pricehat <- data.frame(price01, price02, price03, price04, price05)
pricehat$mu <- rowMeans(pricehat)

##      This tells us a little bit about the estimate.
boxplot(pricehat$mu, horizontal=TRUE, main="Worldwide cash withdrawal prices \n Boxplot of country estimates")
summary(pricehat)

```



$\square$

---

Ben Mazzotta is a postdoc at [IBGC](http://fletcher.tufts.edu/IBGC) and part of the [Cost of Cash](http://fletcher.tufts.edu/CostofCash) team.