---
title:  Multiple imputation for missing data on cash prices.
author: Ben Mazzotta
affiliation:  Cost of Cash Global
date:  10/1/2014
output: pdf_document
---

## Goals

1. Begin with remittances dataset and financial inclusion covariates.
2. Impute prices for cashout transactions where unavailable.
3. Impute rates for caashout transactions where unavailable.
4. Model national costs of cashout transactions from combined dataset, i.e., imputed and reported.


```{r preliminaries, comment=F, warning=FALSE, echo=FALSE}
#    Amelia package does imputation
require(Amelia) ; require(data.table) ; require(ggplot2)

#   open the data
load("../data/remittances.Rdata")

#   join financial inclusion to remittances data

```


## Impute prices 

Cashout transactions are poorly reported. Fewer than half of countries in the remittance prices worldwide dataset have average prices for a cashout transaction. None of these are rich countries.

Cashout transactions are the basis for our estimates of cash fees. Without them we are dead in the water. So what are those fees? We need to impute the data.

Covariates of cashout fees

* financial access
* bank density per capita
* remittances inbound
* gdp per capita
* PPP price level ratio

```{r imputeprices}

```


#     1. Open the datasets

#     2.  