---
title: "Problem: The Tax Gap"
author: "Ben Mazzotta"
date: "Friday, October 24, 2014"
output: pdf_document
header-includes:
  - \usepackage{times}
  - \usepackage{euler}
---

## Tax Gap

The tax gap is defined as the difference between the true tax burden of a country and its actual tax revenue. In the United States, a quintennial estimate is released at the IRS website using audit data. The most recent applies to tax year 2006, see [The Tax Gap](http://www.irs.gov/uac/The- Tax-Gap).  Needless to say no comparable international study exists.

## Instructions

1. Collect average effective tax rates for as many countries as possible.
2. Collect informal economy to GDP ratios for as many countries as possible.
3. Join the data.
4. Summarize the availability of data.
5. Create a scatterplot of informality and average effective tax rates.
6. Save the joined data in each of the following formats: CSV, DTA, RDA. If you have cloned this repository, make sure to save the data in the `./data` directory


```{r demonstrateWDI, width=30}
##  This is how to pull statistics from World Development Indicators into R.
##      A similar interface exists for Stata. 

#       Load the library
require("WDI")
#       Query variables on taxes
WDIsearch("Tax revenue")
#       Jot down the name of the variable
taxvar <- WDIsearch("Tax revenue")[4][1]  ## Note the subscripts in square brackets
#       Query for all countries, three recent years
tax <- WDI(indicator=taxvar, country="all", start=2010, end=2012, extra=TRUE)
#       Filter for country observations
tax <- subset(tax, region!="aggregates")
names(tax)[grep("GC.", names(tax))] <- "tax.gdp"

rm(taxvar)
summary(tax)

#       Now for informal economy. 
WDIsearch("informal sector")
cat("Rats.")

```

Since the World Bank has no readily available measures of employment in the informal sector, here is the ILO website that measures the same.

http://laborsta.ilo.org/informal_economy_E.html

Good instructions on reading data formats are available with Google searches. I recommend bookmarking

* [UCLA pages](http://www.ats.ucla.edu/stat/r/) on learning R
* [Quick-R](http:statmethods.net)
* StackExchange

## Merge data in R.

I will give an example readily available in WDI, rather than showing you how to import ILO data here. 

```{r learntomerge, fig.width=10, fig.height=6}
require(ggplot2)

##      Pull income per capita
incvar <- WDIsearch("GNI per capita")[2][1]
inccap <- WDI(incvar, country="all", start=2012, end=2012)
names(inccap)[grep("NY.", names(inccap))] <- "inccap"
names(inccap); names(tax)

##      This is the merge command
df <- merge(inccap, tax, by=c("country","iso2c"))
names(df)

##      Keep only the useful variables
df <- df[,c("iso2c","inccap","tax.gdp","region","income","lending","year.y")]
###     Rename the awkward "year.y"
names(df)[names(df) %in% "year.y"] <- "year"
summary(df)
str(df)

##      Simple plot with the base functions
plot(df$tax.gdp, df$inccap, log="y")

##      Fancier GGPLOT2 package
qplot(tax.gdp, inccap, data=subset(df, year==2012), geom="text", label=iso2c, log="y")
```

## Next steps

After you've done that, consider what you would do about incomplete data. How could you model missing data?


---

Ben Mazzotta is a [postdoc](http://fletcher.tufts.edu/costofcash/) at [IBGC](http://fletcher.tufts.edu/ibgc). [Fork me on Github](http://www.github.com/benmazzotta). Check out the CCglobal repository Fletcher.